
%-------------------------------------------------------------------------
%	QUOTATION PAGE
%-------------------------------------------------------------------------
%\quotepage{Matt Smith as \emph{The Doctor}, written by Matthew Graham}
%{
%	I am and always will be the optimist, the hoper of far-flung hopes and the
%	dreamer of \newline improbable dreams
%}

%-------------------------------------------------------------------------
%	DEDICATORY
%-------------------------------------------------------------------------

%\begin{dedicatory}
%	Dedicated to (optional) 
%\end{dedicatory}

%-------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS PAGE
%-------------------------------------------------------------------------
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}
\begin{acknowledgements}

I would like to express my gratitude to my supervisor, Prof. João Pedro Pedroso, for all the availability, guidance, and support throughout this year. The advise and feedback always helped me stay on track.

To my parents, thank you for always believing in me and being my biggest supporters. Your encouragement gave me the strength to overcome the most challenging moments.

To my brother, thank you for the laughs, distractions, and for always being there when I needed it.

To my boyfriend, I am grateful for the endless support, motivation, and love. Thanks for listening, reminding me to take breaks, and always being willing to help.

Finally, to my university friends, thank you for sharing this journey with me. You made the experience lighter, more enjoyable, and truly unforgettable.

\end{acknowledgements}
\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}
%\addvspacetoc{0.3cm} % Add a gap in the Contents, for aesthetics


%-------------------------------------------------------------------------
%	ABSTRACT PAGE (PORTUGUESE)
%-------------------------------------------------------------------------
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}
\begin{abstract}[
	thesistitle={Uma ferramenta interativa de apoio à elaboração de horários universitários},
	title={Resumo},
	degree={Mestrado em Engenharia de Redes e Sistemas Informáticos},
	nameconnector={por},
        keywordsname={Palavras-chave},
        keywords={\textit{University Course Timetabling Problem}, \textit{Curriculum-based Course Timetabling}, \textit{Monte Carlo Tree Search}, \textit{Hill Climbing, Diving}, ITC-2007}]
\begin{otherlanguage}{portuguese}

A geração de horários para escolas e universidades é um problema de otimização combinatória NP-difícil, caracterizado por um espaço de pesquisa vasto e por restrições. Na Faculdade de Ciências da Universidade do Porto (FCUP), a elaboração dos horários a cada semestre para milhares de estudantes, dezenas de currículos e centenas de docentes é, atualmente, um processo manual, demorado e sujeito a erros.

Esta dissertação aborda a variante \textit{Curriculum-Based Course Timetabling} (CB-CTT) do \textit{University Course Timetabling Problem}, apresentando uma abordagem híbrida inovadora que combina \textit{Monte Carlo Tree Search} (MCTS) com \textit{Hill Climbing} (HC) e \textit{diving}. O MCTS conduz a pesquisa global, simulando e avaliando múltiplas atribuições de evento, período e sala, priorizando os eventos mais restritos e expandindo os ramos mais promissores. Sempre que o MCTS encontra, durante a fase de simulação, um horário viável melhorado, o HC assume o controlo e aplica seis movimentos de vizinhança para refinar iterativamente a solução, enquanto preserva a sua viabilidade. A estratégia de diving foi introduzida para melhorar a efetividade do MCTS. Essa estratégia aprofunda soluções parciais promissoras, permitindo que o algoritmo se foque em regiões de maior qualidade no espaço de pesquisa.

Testes com as instâncias do terceiro \textit{track} da ITC-2007 demonstram que a abordagem híbrida encontra consistentemente horários viáveis. No entanto, mesmo com limites de tempo alargados, a abordagem ainda não supera os melhores resultados publicados no que diz respeito à minimização de \textit{soft constraints}.

Apesar de várias melhorias e adaptações ao algoritmo MCTS padrão, continuam a existir desafios. Ainda assim, a abordagem híbrida proposta revela um grande potencial e constitui uma base sólida para investigação futura e aplicação prática em contextos institucionais de criação de horários, como é o caso da FCUP.

\end{otherlanguage}
\end{abstract}
\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}
%-------------------------------------------------------------------------
%	ABSTRACT PAGE
%-------------------------------------------------------------------------
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}
\begin{abstract}

Timetable generation for schools and universities is an NP-hard combinatorial optimization problem with a vast search space and constraints. At the Faculty of Sciences of the University of Porto (FCUP), preparing semester schedules for thousands of students, dozens of curricula, and hundreds of lecturers is currently manual, time-consuming, and prone to errors.

This dissertation addresses the Curriculum-Based Course Timetabling (CB-CTT) variant of the University Course Timetabling Problem by presenting a novel hybrid approach that combines Monte Carlo Tree Search (MCTS) with Hill Climbing (HC) and a diving strategy. MCTS drives the global search, simulating and evaluating multiple event, period, and room assignments, prioritizing the most constrained events and expanding the most promising branches. Whenever MCTS obtains an improved feasible timetable during simulation, HC takes over and applies six neighborhood moves to iteratively improve the solution while preserving the feasibility. The diving strategy was introduced to further enhance the MCTS effectiveness. This strategy deepens promising partial solutions, allowing the algorithm to focus on higher-quality regions of the search space.

Tests on the ITC-2007 track 3 benchmark instances show that the hybrid approach consistently finds feasible timetables. However, even with extended time limits, the approach does not yet outperform best published results in terms of minimizing soft constraint violations.

Despite some improvements and changes to the standard MCTS algorithm, challenges remain. Nevertheless, the proposed hybrid approach demonstrates strong potential and a solid foundation for further research and practical adoption in institutional timetabling contexts like FCUP. 

\end{abstract}
\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}

%-------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES
%-------------------------------------------------------------------------

\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}

\tableofcontents % Write out the Table of Contents

\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}
\addvspacetoc{0.3cm}

%\listoftables % Write out the List of Tables

\listoffigures % Write out the List of Figures



%\addvspacetoc{0.3cm}

%-------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%-------------------------------------------------------------------------

%\begin{listofcontants}
%	\const{My little ponny test of magical rainbow}{$mn/mp$}
%    {$2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$}
%   \const{Vaccuum permeability test of magical rainbow for a specific case of
%   condensed matter physics}
%   {$\epsilon_0$}{$2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$}
%	\const{Speed of Light test of magical rainbow}{$c$}
%    {$2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$}
%\end{listofcontants}


%-------------------------------------------------------------------------
%	SYMBOLS
%-------------------------------------------------------------------------

%\begin{listofsymbols}
%	\symb{$F_{\mu\nu}$}{Maxwell tensor}{F}
%	\symb{$a$}{distance}{m}
%	\\
%	\symb{$\omega$}{angular frequency}{rads$^{-1}$}
%\end{listofsymbols}


%-------------------------------------------------------------------------
%	NOTATION
%-------------------------------------------------------------------------

% \newcommand\notationname{Notation and Conventions}
% \addtotoc{\notationname}
% \fancyhead[LO]{\textsc{\notationname}}

% \input{Notation}



%-------------------------------------------------------------------------
%	ABBREVIATIONS
%-------------------------------------------------------------------------

\newacronym{fcup}{FCUP}{Faculty of Sciences of the University of Porto}
\newacronym{ucttp}{UCTTP}{University Course Timetabling Problem}
\newacronym{mcts}{MCTS}{Monte Carlo Tree Search}
\newacronym{itc}{ITC}{International Timetabling Competition}
\newacronym{itc-2007}{ITC-2007}{Second International Timetabling Competition}
\newacronym{patat}{PATAT}{International Conference on the Practice and Theory of Automated Timetabling}
\newacronym{uct}{UCT}{Upper Confidence Bounds for Trees}
\newacronym{cb-ctt}{CB-CTT}{Curriculum-based Course Timetabling}
\newacronym{pe-ctt}{PE-CTT}{Post-Enrollment Course Timetabling}
\newacronym{hc}{HC}{Hill Climbing}
\newacronym{ts}{TS}{Tabu Search}
\newacronym{sa}{SA}{Simulated Annealing}
\newacronym[plural=COPs, firstplural=Combinatorial Optimization Problems (COPs)]{cop}{COP}{Combinatorial Optimization Problem}

\printglossary[type=\acronymtype,title={List of Abbreviations}]
